---
title: "Inferring ancestral founders from simulated HIV-1 sequences"
output:
html_document:
theme: cerulean
---

HIV-1 evolution was simulated for 50,000 generations, sampling 100 sequences from each generation.

From this single evolutionary lineage, two generations were
 sampled to simulate a transmission hierarchy.
 For example, 10 sequences were selected from generation 400, and 10
 sequences were selected from generation 1000.  These sequences were
 grouped into a mono-phyletic taxa, forming two main branches for a
 phylogenetic tree. BEAST was used to infer sequences at the root of
 this tree.

 One branch was kept constant, selected from generation 400, while
 the other branch was selected from generations, 100, 500, 1000, and
 10,0000 of the simulated lineage.  These datasets were used to infer ancestral sequeces for strict and relaxed molecular clock models, resulting in eight BEAST runs in total.
 

```{r echo=FALSE, results='hide'}
library(ggplot2)
library(scales)
library(plyr)
library(ape)

tryCatch(
{
    df <- read.csv("../sims/distances.csv", comment="#")
},
error=function(e) {
    stop('Cannot open "../sims/distances.csv"')
})


# Show the tree corresponding to the first entry in the distances data frame.
fn <- with(df[which(df$clock=="relaxed")[1:4],,drop=FALSE], { sprintf("../sims/runs/%d/%d/%s/mcc.tree", left, right, clock) })
message(sprintf("reading \"%s\"", fn))
lapply(fn, function(fn) {
tree <- read.nexus(fn)  # nexus format
plot(tree,cex=0.7,root.edge = TRUE, direction = "rightwards", main=fn)
box(lwd = 2)

})


```

Weighted distance from actual founding sequence of consensus sequence
at MRCA of patient2, weighted by tree posterior probability.
```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE, results='hide'}
    p <- ggplot(df, aes(x=right, y=exp.wdist)) +
         geom_line() +
         geom_point() +
              facet_grid(clock~., scales="free_y", space="free") +
         ylab('Distance (normalized mismatches\nweighted by posterior)') +
         xlab('Right-hand branch length (generations)')
    p <- p + ggtitle("Weighted distance\n- inferred sequence to actual founder -\nby right-hand branch length") + 
         theme(plot.title = element_text(lineheight=.8, face="bold"))

print(p)


```



Weighted distance from actual founding sequence of consensus sequence
at MRCA of patient2, weighted by tree posterior probability.
```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE, results='hide'}
    p <- ggplot(df, aes(x=right, y=exp.wdist)) +
         geom_line() +
         geom_point() +
              facet_grid(clock~., scales="free_y", space="free") +
         ylab('Distance (normalized mismatches\nweighted by posterior)') +
         xlab('Right-hand branch length (generations)')
    p <- p + ggtitle("Weighted distance\n- inferred sequence to actual founder -\nby right-hand branch length") + 
         theme(plot.title = element_text(lineheight=.8, face="bold"))

print(p)


```


Log weighted distance of consensus sequence at MRCA from actual founding sequence, weighted by tree posterior probability.
```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE, results='hide'}
    p <- ggplot(df, aes(x=right, y=wdist)) +
         geom_line() +
             geom_point() +

              facet_grid(clock~., scales="free_y", space="free") +
         ylab('Log-Distance (normalized mismatches\nweighted by posterior)') +
         xlab('Right-hand branch length (generations)')
    p <- p + ggtitle("Log-Weighted distance\n- inferred sequence to actual founder -\nby right-hand branch length") + 
         theme(plot.title = element_text(lineheight=.8, face="bold"))

print(p)


```


Distance of consensus sequence at MRCA from actual founding sequence, segregated by molecular clock model.
```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE, results='hide'}
    p <- ggplot(df, aes(x=clock, y=mean)) +
         geom_boxplot() +
         ylab('Mean distance (mismatches)') +
         xlab('molecular clock')
    p <- p + ggtitle("Average distance\n- inferred sequence to actual founder -\nby molecular clock") + 
         theme(plot.title = element_text(lineheight=.8, face="bold"))

print(p)


```
