---
title: "Inferring ancestral founders from simulated HIV-1 sequences"
output: html_document
theme: cerulean
runtime: shiny
---

Patient1 is a simulated HIV-1 lineage covering 10,000 generations.
At various points in the patient 1 lineage, a single sequence was
sampled to use as the founder of a new lineage, "patient2".


HIV-1 evolution was simulated for 10,000 generations, sampling 100 sequences from each generation.

From this single evolutionary lineage, two generations were
 sampled to simulate a transmission hierarchy.
 For example, 10 sequences were selected from generation 400, and 10
 sequences were selected from generation 1000.  These sequences were
 grouped into a mono-phyletic taxa, forming two main branches for a
 phylogenetic tree. BEAST was used to infer sequences at the root of
 this tree.

 One branch was kept constant, selected from generation 400, while
 the other branch was selected from generations, 100, 500, 1000, and
 10,0000 of the simulated lineage.  These datasets were used to infer ancestral sequeces for strict and relaxed molecular clock models, resulting in eight BEAST runs in total.
 

```{r echo=FALSE}


# To run this dynamic document,
# library(shiny)
# run("dynalysis.Rmd", shiny_args=c(host="140.107.116.67", launch.browser=FALSE))

library(ggplot2)
library(scales)
library(plyr)
library(ape)
library(ggvis)
library(Biostrings)
library(shiny)

# hack to get around missing title function in ggvis
# http://stackoverflow.com/a/25030002/1135316
add_title <- function(vis, ..., title = "Plot Title") 
{
  add_axis(vis, "x") %>% 
  add_axis("x", orient = "top", ticks = 0, title = title,
           properties = axis_props(
             axis = list(stroke = "white"),
             labels = list(fontSize = 0),
             title=list(fontSize = 22)
           ), ...)
}


tryCatch(
{
    df <- read.csv("../sims/distances.csv", comment="#")
},
error=function(e) {
    stop('Cannot open "../sims/distances.csv"')
})

# annotate each row with a unique value that wil be used to track values as graphs change.
df$id <- 1:nrow(df)


```

```{r, fig.width=4, fig.height=5, echo = FALSE}

## The input controls and data frame columns are named as follows:
##     Tinf.p1 - time since infection, patient 1
##     Ttrans - time of transmission, along patient 1 timeline
##     Tinf.p2 - time since infection, patient 2

inputPanel(
    selectInput("tree.Tinf.p1", "Patient-1 time since infection:",
                choices = sort(unique(df$Tinf.p1))),
    selectInput("tree.Ttrans", "Transmission time:",
                choices = sort(unique(df$Ttrans))),
    selectInput("tree.Tinf.p2", "Patient-2 time since infection:",
                choices = sort(unique(df$Tinf.p2)))
    )

observe({
    choices = sort(unique( df[df$Tinf.p1 == input$tree.Tinf.p1,]$Ttrans ))
        updateSelectInput(session, "tree.Ttrans", choices = choices)

})

observe({
    choices = sort(unique( df[df$Tinf.p1 == input$tree.Tinf.p1 & df$Ttrans == input$tree.Ttrans,]$Tinf.p2 ))
        updateSelectInput(session, "tree.Tinf.p2", choices = choices)

})



```

```{r, fig.width=4, fig.height=5, echo = FALSE}

renderPlot({
    message("in renderplot")
    fn <- sprintf("../sims/runs/%s/%s/%s/relaxed/mcc.tree", input$tree.Tinf.p1, input$tree.Ttrans, input$tree.Tinf.p2) 
    tree <- read.nexus(fn)  # nexus format
    plot(tree,root.edge = TRUE, direction = "rightwards", main=fn, cex.main=3)
    add.scale.bar(cex = 0.7, font = 2, col = "red")

    box(lwd = 2)
  })



```

```{r, , fig.width=4, fig.height=5, echo = FALSE}

distance.data <- reactive({
    if (is.null(input$tree.Tinf.p1) || is.null(input$tree.Ttrans) || is.null(input$tree.Tinf.p2) )
        return(NULL)
    log.file <- sprintf("../sims/runs/%s/%s/%s/relaxed/ancestralSequences.log", input$tree.Tinf.p1, input$tree.Ttrans, input$tree.Tinf.p2)
    founder.file <- sprintf("../sims/runs/%s/%s/founder.fa", input$tree.Tinf.p1, input$tree.Ttrans)
    if (!all(file.exists(c(log.file, founder.file)))) {
        message("returning NULL from distance.data")
        return(NULL)
    }
    df <- read.table(log.file, comment="#", sep="\t", header=TRUE, stringsAsFactors=F)
    founder <- readDNAStringSet(founder.file, use.names=FALSE)
    nmismatch <- function(x,y) {
        foo <- function(x,y) {
            sum(mapply('!=', strsplit(as.character(x),split=''), strsplit(as.character(y),split='')))
        }
        mapply(foo, x, y)
    }
    nmismatch(founder, df$trait)
})

# Plot how many times each distance measure is seen in a particular configuration.
#
# Each time beast runs, it generates many trees and many inferred ancestral sequences.
# If we measure the distance from each of these sequences to the founder we can see how diverse a
# sampling of ancestor sequences beast is generating.
#


ggvisOutput('histdistance')
invisible(reactive({
    distance = distance.data()
    if (is.null(distance)) {
        data.frame(distance=1:10) %>% ggvis(x = ~distance ) %>% layer_histograms() %>%
             add_title(title = "Frequency of distance measures") 
    } else {
        data.frame(distance) %>% ggvis(x = ~distance ) %>% layer_histograms() %>%
            add_axis("y",title_offset = 50, properties = axis_props(title=list(fontSize = 18))) %>%
            add_axis("x",title_offset = 50, properties = axis_props(title=list(fontSize = 18))) %>%
                add_title(title = sprintf("Frequency of distance measures %s/%s/%s",
                                          input$tree.Tinf.p1, input$tree.Ttrans, input$tree.Tinf.p2))

    } 
    })   %>% bind_shiny('histdistance'))

```


Weighted distance from actual founding sequence of consensus sequence
at MRCA of patient2, weighted by tree posterior probability.

```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE}

inputPanel(
    selectInput("Tinf.p1", "Patient-1 time since infection:",
                choices = c("-all-",sort(unique(df$Tinf.p1)))),
    selectInput("Ttrans", "Transmission time:",
                choices = c("-all-",sort(unique(df$Ttrans)))),
    selectInput("Tinf.p2", "Patient-2 time since infection:",
                choices = c("-all-",sort(unique(df$Tinf.p2))))
    ) # end-inputPanel

observe({
    if (!is.null(input$Tinf.p1) &&  input$Tinf.p1!= "-all-") 
        choices = c("-all-",sort(unique( df[df$Tinf.p1 == input$Tinf.p1,]$Ttrans )))
    else
        choices = c("-all-", sort(unique( df$Ttrans )))

    updateSelectInput(session, "Ttrans", choices = choices)

})

observe({
    if (!is.null(input$Ttrans) && input$Ttrans != "-all-") {
        if (!is.null(input$Tinf.p1) && input$Tinf.p1 != "-all-") {
            choices = c("-all-",sort(unique( df[df$Tinf.p1 == input$Tinf.p1 & df$Ttrans == input$Ttrans,]$Tinf.p2 )))
        } else {
            choices = c("-all-",sort(unique( df[df$Ttrans == input$Ttrans,]$Tinf.p2 )))
        }
    } else {
        choices = c("-all-",sort(unique( df$Tinf.p2 )))
    }
    updateSelectInput(session, "Tinf.p2", choices = choices)
})


xyzzy <- reactive({
    data <- df
    if (!is.null(input$Tinf.p1) && input$Tinf.p1 != "-all-") {
        data = data[data$Tinf.p1 == input$Tinf.p1,]
    } 
    if (!is.null(input$Tinf.p2) && input$Tinf.p2 != "-all-") {
        data = data[data$Tinf.p2 == as.numeric(input$Tinf.p2),]
    } 
    if (!is.null(input$Ttrans) && input$Ttrans != "-all-") {
        data = data[data$Ttrans == as.numeric(input$Ttrans),]
    } 
    data
})


```



```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE}

# given the data under a point, return an HTML string for display under the mouse
tooltipString <- function(x) {
    if (is.null(x)) return(NULL)
    print(x)
    row <- df[df$id == x$id,]
    
    print(row)
    paste0("P1 infected period: ", row[['Tinf.p1']], "<br />",
           "   transmission time: ", row[['Ttrans']], "<br />",
           "P2 infected period: ", row[['Tinf.p2']])
    
}

ggvisOutput('scatterplot')
invisible(reactive({
    data <- xyzzy()
    data %>% ggvis(~ Tinf.p1 - Ttrans, ~Tinf.p2, key:= ~id, size = ~dist.mean ) %>% layer_points()  %>%
    add_axis("x", title = "Patient-1 time since transmission(generations)") %>%
    add_axis("y", title = "Patient-2 time since infection(generations)", title_offset = 50) %>%
    scale_numeric("x", domain = c(min(df$Tinf.p1), max(df$Tinf.p1)), nice = FALSE) %>%
    # scale_numeric("y", domain = c(min(df$Tinf.p2), max(df$Tinf.p2)), nice = FALSE) %>%
    add_title(title = "Average Distance from Founder")
    # add_tooltip(tooltipString, "hover") 
   })   %>% bind_shiny('scatterplot'))



```




Mean distance of consensus sequence at MRCA from actual founding sequence, segregated by molecular clock model.
```{r eval=TRUE, fig.width=6, fig.height=5, echo=FALSE, results='hide'}

    p <- ggplot(df, aes(x=clock, y=mismatch.mean)) +
         geom_boxplot() +
         ylab('Mean distance (mismatches)') +
         xlab('molecular clock')
    p <- p + ggtitle("Average distance\n- inferred sequence to actual founder -\nby molecular clock") + 
         theme(plot.title = element_text(lineheight=.8, face="bold"))

print(p)


```

